Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2013-11-28 masters/body/_region_.tex
gain comp

\message{ !name(masters.tex) !offset(135) }
\subsection{Geometry Partitioning}
\label{sec:geometryPartioning}
In order to effectively texture a 3D model, we first divide the model
into a set of regions, each to be textured independently. The purpose
of this is to ensure texture continuity along large uniform areas,
while allowing texture boundaries to fall along natural environmental
boundaries. Because textures may not match at such boundaries where
two regions meet, it is important to minimize the visibility of
boundaries. This is done by encouraging region boundaries to occur at
large sharp corners, where texture continuity is less important.

When working with low-resolution models, where surfaces generally
correspond to large planar features such as walls, ceilings, or
floors, and often meet at right angles, each planar surface in the
model is directly treated as a distinct region and textured
separately.

When working with a high-resolution model however, geometry often
represents large features as well as small details found in furniture,
plants, and other miscellaneous objects. For instance, in Figure
\ref{fig:deskGeom}, the geometry modeling the desk and binders
consists of many small triangles at various orientations. Texturing
each of these triangles independently makes little sense, as the
textures applied to each should match up well with the textures of its
neighbors in order to avoid any breaks in the overall image of the
desk and binders. On the other hand, the textures for the triangles
belonging to the wall behind need not match up with the textures on
the desk and binders, and so each group of triangles should be
designated to separate regions. This region grouping is performed by designating all contiguous coplanar groups of triangles as different regions. Any region that is less than $1 m^2$ in size is joined to its largest neighboring region, such that the angle between them is under $90^{\circ}$.

Because a texture is 2D, and the region to be textured may not be
flat, a mapping needs to be defined. We first fit a plane to each
region for the purposes of texturing. This can be accomplished via a
least squares plane-fitting through each triangle vertex, though in
practice we simply use the plane equation of the largest contiguous
section of coplanar triangles, as most regions are characterized by
small outcroppings from a largely planar surface, such as objects on a
table or features protruding from a wall. The 3d vertices within a
region are then projected onto its calculated plane, and the
rectangular bounding box for these projected vertices is
calculated. This bounding box corresponds to the 2D image that will be
created to texture the region.

Of course, the area encompassed by the bounding box is larger than the
corresponding 2d area of the projected geometry, which itself is not
as accurate as the original 3d geometry. Thus, we also create a k-d
tree consisting of the triangle primitives from the original
mesh. Though we generate texture for the 2d bounding box surface, we
will reference the original 3D geometry to ensure accuracy when
texturing.

\message{ !name(masters.tex) !offset(944) }

>>>>

\message{ !name(masters.tex) !offset(890) }
\subsection{Exposure Compensation}
\label{sec:exposureCompensation}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3.3in]{exposureDiff1.png}}
  \subfloat[][]{\includegraphics[width=3.3in]{exposureDiff2.png}}
  \caption{(a) A ceiling texture composed of images taken with varying
    exposures, with significant brightness differences. (b) Exposure
    compensation applied to the same set of images from (a) by
    applying computed gains to each image}
  \label{fig:exposureDiff}
\end{figure}

Before blending images together, exposure compensation is applied to
equalize brightness among neighboring images. For the images in this
report, cameras are set to have automatic exposure, which means that
images of the same object may have different brightness levels. While
successful image alignment and blending can reduce sharp seams between
adjacent images, there may still be noticeable brightness gradients
between images, particularly in areas near light sources. This can be
seen in Figure \ref{fig:exposureDiff}(a), where the ceiling texture
has patches of clearly differing brightness, most noticeably around
light sources. To diminish this effect, a gain can be computed and
applied to each image, with the effect of linearly scaling the
brightness of an image. The goal is to compute a gain for each image,
such that the brightness of an area is consistent across image
boundaries.

Similar to the image location adjustment procedure in Section
\ref{sec:robustSIFTFeatureMatching}, exposure compensation can be
neatly formed as a least-squares optimization problem with pair-wise
observations. In this case, observations do not need to be weighted,
and so the formulation is $\textrm{min}_{\vec{\beta}} ||(A \vec{\beta}
- \vec{\gamma})||_2^2 $. An example setup for 3 images, each with 2
overlapping pixels is shown below.
\[
A =
\begin{pmatrix}
  P_{11} & -P_{12} & 0\\
  P_{21} & -P_{12} & 0\\
  0 & P_{32} & -P_{33}\\
  0 & P_{42} & -P_{43}\\
  1 & 1 & 1\\

\end{pmatrix}\quad
\vec{\beta} =
\begin{pmatrix}
  G_1, \\ G_2, \\ G_3
\end{pmatrix}
\vec{\gamma} =
\begin{pmatrix}
  0, \\ 0, \\ 0, \\ 0, \\ 3
\end{pmatrix}
\]

In this problem, $\beta$ is the vector of gains $G_i$ to be solved
for.  The observations, represented by $A$ and $\gamma$, correspond to
equalizing the brightness for each pixel present in two images. For
instance, for pixel $P_x$ which is present in two overlapping images
$I_i$ and $I_j$, the goal is to find gains $G_i$ and $G_j$ such that
$G_iP_{xi} - G_jP_{xj} = 0$, where $P_{xy}$ is the value of pixel
$P_x$ in image $I_y$. Thus, $\gamma$ contains all zeros, while each
row of $A$ contains a $P_i$ value in the $ith$ column and a $-P_j$
value in the $jth$ column, and zeros elsewhere. The number of
observations is thus equal to the number of pixels that occur in two
distinct images.  This problem is not constrained, as it allows all
computed gains to be scaled by any amount. To keep gains at reasonable
values, a final observation is added: $\Sigma G_i = N$, where $N$ is
the number of images. This has the effect of averaging all gains to 1.

The result of computing gains and equalizing exposure can be seen for
the same ceiling location in Figure \ref{fig:exposureDiff}(b). The
effect is clear, as dark/bright patches are no longer present, and the
overall brightness of the image has not changed significantly.

\message{ !name(masters.tex) !offset(1248) }

### 2013-11-28 masters/body/masters.tex
gain comp

\subsection{Blending}
\label{sec:blending}
We now apply a blending procedure to the texturing methods in Sections
\ref{sec:mappingWithCaching} and \ref{sec:shortestPath}. Although the
image alignment steps and image selection in both methods attempt to
minimize all mismatches between images, there are occasional
unavoidable discontinuities in the final texture due to different
lighting conditions or inaccuracies in model geometry. These can
however be treated and smoothed over by applying alpha blending over
image seams.  Whether the units to be blended are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as there is a guaranteed overlap
between units to blend over.

>>>>

\subsection{Exposure Compensation}
\label{sec:exposureCompensation}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3.3in]{exposureDiff1.png}}
  \subfloat[][]{\includegraphics[width=3.3in]{exposureDiff2.png}}
  \caption{(a) A ceiling texture composed of images taken with varying
    exposures, with significant brightness differences. (b) Exposure
    compensation applied to the same set of images from (a) by
    applying computed gains to each image}
  \label{fig:exposureDiff}
\end{figure}

Before blending images together, exposure compensation is applied to
equalize brightness among neighboring images. For the images in this
report, cameras are set to have automatic exposure, which means that
images of the same object may have different brightness levels. While
successful image alignment and blending can reduce sharp seams between
adjacent images, there may still be noticeable brightness gradients
between images, particularly in areas near light sources. This can be
seen in Figure \ref{fig:exposureDiff}(a), where the ceiling texture
has patches of clearly differing brightness. To diminish this effect,
a gain can be computed and applied to each image, with the effect of
linearly scaling the brightness of each image. The goal is to compute
a gain for each image, such that the brightness of an area is
consistent across image boundaries.

Similar to the image position refinement procedure in Section
\ref{sec:robustSIFTFeatureMatching}, exposure compensation is
performed simulatenously across all images present in a single
region. As before, this is solved as a least-squares optimization
problem with pair-wise observations. In this case, observations do not
need to be weighted, and so the formulation is
$\textrm{min}_{\vec{\beta}} ||(A \vec{\beta} - \vec{\gamma})||_2^2
$. An example setup for 3 images, each with 2 overlapping pixels, is
shown below.
\[
A =
\begin{pmatrix}
  P_{11} & -P_{12} & 0\\
  P_{21} & -P_{12} & 0\\
  0 & P_{32} & -P_{33}\\
  0 & P_{42} & -P_{43}\\
  1 & 1 & 1\\

\end{pmatrix}\quad
\vec{\beta} =
\begin{pmatrix}
  G_1, \\ G_2, \\ G_3
\end{pmatrix}
\vec{\gamma} =
\begin{pmatrix}
  0, \\ 0, \\ 0, \\ 0, \\ 3
\end{pmatrix}
\]

In this problem, $\beta$ is the vector of gains $G_i$ to be solved
for.  The observations, represented by $A$ and $\gamma$, correspond to
equalizing the brightness for each pixel present in two images. For
instance, for pixel $P_x$ which is present in two overlapping images
$I_i$ and $I_j$, the goal is to find gains $G_i$ and $G_j$ such that
$G_iP_{xi} - G_jP_{xj} = 0$, where $P_{xy}$ is the value of pixel
$P_x$ in image $I_y$. Thus, $\gamma$ contains all zeros, while each
row of $A$ contains a $P_{xi}$ value in the $ith$ column and a
$-P_{xj}$ value in the $jth$ column, and zeros elsewhere. This problem
is not constrained, as it allows all computed gains to be scaled by
any amount. To keep gains at reasonable values, a final observation is
added: $\Sigma G_i = N$, where $N$ is the number of images. This has
the effect of setting the average of all gains to 1, and is
represented by adding a row of ones to $A$, and a corresponding $N$
value to $\gamma$.

The result of computing and applying gains for each image can be seen
for the same ceiling location in Figure \ref{fig:exposureDiff}(b). As
compared to Figure \ref{fig:exposureDiff}(a), the effect is clear, as
dark/bright patches are no longer present, and the overall brightness
of the image has not changed significantly.

\subsection{Blending}
\label{sec:blending}
We now apply a blending procedure to the texturing methods in Sections
\ref{sec:mappingWithCaching} and \ref{sec:shortestPath}. Although the
image alignment and selection steps in both methods attempt to
minimize all mismatches between images, and exposure compensation
minimizes brightness differences, there are occasional unavoidable
discontinuities in the final texture due to parallax effects or
inaccuracies in model geometry. These can however be treated and
smoothed over by applying alpha blending over image seams.  Whether
the units to be blended are rectangularly-cropped images or
rectangular tiles, we can apply the same blending procedure, as long
as there is a guaranteed overlap between units to blend over.
